Leadership across cultures and contexts: an exploratory analysis

### 2020-05-13 leadership_across_cultures_contexts.Rmd
Added non-WEIRD vs. WEIRD paragraph in conclusion.

While there is evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies. There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms. For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies. It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].

>>>>

Our results also suggest both broad continutites and contrasts between leadership among nonindustrial, non-WEIRD populations sampled here and he WEIRD samples common in managerial studies. Our organize and strategize features of leader functions reflect the @wofford_factor_1970 *order and group achievement* and *personal interaction* factors, respectively. Also, Woffords *personal enhancement* and *dynamic and achievement* factors are similar to leadership dimensions captured by our compentencies feature. Our results, however, emphasize the prosociality of leaders in non-WEIRD settings over an emphasis on self-presevation, relative to WEIRD leaders (e.g., Woffords *security and maintenace* factor). We also identified conformity to social norms as a feature of human leadership across cultures, similar to predictions and findings drawing on Culturally Endorsed Implicit Leadership Theory by managerial scholars working with WEIRD populations [@behrendt_integrative_2017]. Our results on the functions of leaders across cultures and contexts also corresponds to the *task-oriented* vs. *relation-oriened* framework of managerial beahviors (i.e., our Managemebt vs. Prosociality features of leader functions), however, our results do not support the strong emphasis on charismastic leadership found by the GLOBE CEO study [@house2013strategic] (although charism was documented as a quality of leaders in over 20% of cultures and oratory skills in over 50%). See Figure \@ref(fig:cluser).

While there is cross-cultural evidence of a dominance style of leadership involving coercive authority, confidence and aggression, as well as a prestige style involving special knowledge, experience, and innovation, we suggest rethinking dichotomous models of leadership styles and focusing on inter-individual variation in specific leadership competencies. There is likely to be substantial overlap in and interaction between physical, spiritual, cognitive, and other prosocial competencies and the degree to which a leader relies on such competencies will be shaped by their individual levels in them, the competencies of potential rivals, as well as their functions and group context, follower preferences, and cultural norms. For example, widespread preferences for egalitarianism and social equality are likely strong drivers of followers preferences for leaders across diverse societies [e.g., @cheng_dominance_2019; @price_evolution_2014], but shamanism and ritual based contexts represent a social domain in which ideologies support a style of leadership fusing both dominance and prestige-based competencies. It seems evident that cultural forces and ideologies play important roles in shaping leadership styles and follower preferences including among WEIRD populations and in national political movements around the world [e.g., @mcdermott_leadership_2018; @inglehart_trump_2016].
